---
title: "Text Mining with R"
output:
  html_document:
    df_print: paged
---

R is a programming language with a number of advanced packages for text mining. Unlike the previous tools, R requires you to use a 'command line interface', meaning you have to type code rather than use an graphical user interface. 

This document won't teach you how to code, but it's intended to give you a flavour of the kind of things you can do with text and a progamming language. If you already know some R, then you can feel free to manipulate the examples in the code below. 

A good place to start to learn R is the free book https://r4ds.had.co.nz/index.html, and if you want to learn how to use R for text mining specifically, https://www.tidytextmining.com/ is an excellent (also free) online book.

This document you're reading is called an R Notebook: it's a special document which contains text *and* code. Anything within the gray block below is called a 'chunk'. On the right-hand side you'll see a little green triangle. Clicking that will run all the code in the chunk. 1 + 1 is very simple computer code which adds those two numbers together. Try clicking on the code. 

```{r}
1+1
```

R runs the code, and displays the output underneath the chunk 

The programming language R is supplemented with what are known as *packages*. These are bundles of additional functions created by third-parties. Some are developed to do very specific tasks, and others are much more general. We need two to do text mining with R: ```tidyverse```, which is a huge collection of tools to generally extend the functionality of R, and ```tidytext```, which was developed specifically to work with text. You need to run the chunk below to load these packages (the functionality won't be available otherwise)

```{r}
library(dplyr)
library(ggplot2)
library(magrittr)
library(tidytext)
library(stringr)
```

Next we need some text. The easiest place to start is with the ```janeaustenr``` package. 

```{r}
library(janeaustenr)
austen_books = janeaustenr::austen_books()
```

This code adds a line and chapter number to each of the books:

```{r}
original_books <- austen_books %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]",
                                           ignore_case = TRUE)))) %>%
  ungroup()
```

Now that we have the text we can analyse it. First run a function called ```unnest_tokens```, which creates a dataset containing each word on a single line. We can then use R functions to sort and count the words:

```{r}
austen_tokens = original_books %>% unnest_tokens(word, text)

```

First, a simple count of the top words:

```{r}
austen_tokens %>% group_by(word) %>% tally() %>% arrange(desc(n)) %>% head(20)
```
Unsurprisingly they are all short function words like 'the', 'and' and so forth. We can remove these, and make a small chart of the results:

```{r}
data("stop_words")

austen_tokens %>% 
  anti_join(stop_words) %>% 
  group_by(word) %>% 
  tally() %>% 
  arrange(desc(n)) %>% 
  head(20) %>% 
  ggplot() + 
  geom_col(aes(x = reorder(word,n), y = n)) + 
  coord_flip() + 
  labs(title = "20 most frequent words in Jane Austen Novels")
```
We can also count the words per novel:

```{r}
austen_tokens %>% group_by(book) %>% tally(name = 'words') 
```
Or chart the appearance of a word in a single book. Try swapping out the book name or word name between the inverted commas below, to count other words.

```{r}
austen_tokens %>% 
  filter(book == 'Pride & Prejudice') %>% 
  filter(word == 'marriage') %>% 
  group_by(chapter) %>% 
  tally() %>% 
  ggplot() + 
  geom_col(aes(chapter, n))
```

